Calculating life expectancy and exploring missingness

### 2016-10-05 Code/1_Explore_Data_v2.Rmd

```{r calculate_life_expectancy}
source("/Users/corinneriddell/Documents/repos/BlackWhiteGap/Code/life_expectancy_functions.R")

@@

make a new dataset that contains information on the gap

```{r}
Black <- dat2 %>% filter(Race2 == "Black" & Age3 == 0) %>% 
                  select(State2, Year3, Sex2, Pop_across_age, life.expectancy.birth) %>%
                  rename(Pop_black = Pop_across_age,  le.birth.black = life.expectancy.birth)

@@

>>>>

```{r calculate_life_expectancy}
source("/Users/corinneriddell/Documents/repos/BlackWhiteMortalityGap/Code/life_expectancy_functions.R")

@@

Make a new dataset that contains information on the gap:

```{r create_dataset_BlackWhite}
Black <- dat2 %>% filter(Race2 == "Black" & Age3 == 0) %>% 
                  select(State2, Year3, Sex2, Pop_across_age, life.expectancy.birth) %>%
                  rename(Pop_black = Pop_across_age,  le.birth.black = life.expectancy.birth)

@@

Decompose the gap by age. That is, for each year-state-sex combination, compare the life table for blacks to whites and decompose the gap in life expectancy at birth by age.

```{r calculate_age_decomp}

#create list of stratum.ids for each race-sex combination
for (rs in unique(levels(dat.clean$RaceSex))) {
  assign(paste0(gsub(" ", "", rs), ".stratum.ids"), 
      dat.clean %>% filter(dat.clean$RaceSex == rs) %>%
        select(StateYearRaceSex, StateYearSex) %>%
        group_by(StateYearRaceSex, StateYearSex) %>%
        filter(row_number() == 1)# %>%
      #  rename(paste0(gsub(" ", "", rs), ".ids") = StateYearRaceSex)
         )

#merge the black-white lists
Female.ids <- merge(WhiteFemale.stratum.ids, BlackFemale.stratum.ids, by = "StateYearSex")
Female.ids <- Female.ids %>% rename(White.StateYearRaceSex = StateYearRaceSex.x, Black.StateYearRaceSex = StateYearRaceSex.y)
Female.ids$Sex2 <- "Female"

Male.ids <- merge(WhiteMale.stratum.ids, BlackMale.stratum.ids, by = "StateYearSex")
Male.ids <- Male.ids %>% rename(White.StateYearRaceSex = StateYearRaceSex.x, Black.StateYearRaceSex = StateYearRaceSex.y)
Male.ids$Sex2 <- "Male"

paired.ids <- rbind(Male.ids, Female.ids)

list.age.decomp.tables <- vector(mode = "list", length = dim(paired.ids)[1])

for(i in 1:dim(paired.ids)[1]) {
  white.index <- which(non.zero.stratums == paired.ids[i, 2])
  black.index <- which(non.zero.stratums == paired.ids[i, 3])

  if(length(white.index) != 0 & length(black.index) != 0) {
    list.age.decomp.tables[[i]] <- le_age_decomp(life.table1 = as.data.frame(list.life.tables[[black.index]]), 
                                                 name.lt1 = "Black", 
                                                 life.table2 = as.data.frame(list.life.tables[[white.index]]), 
                                                 name.lt2 = "White", age.groups = "Age3")
  } else {
    list.age.decomp.tables[[i]] <- "Couldn't compare life tables"
  }
  print(i)
}
